Windowed least-squares model reduction for dynamical systems

### 2019-10-17 section4/spatial_numerical.tex
changed notation in 4.2

Direct approaches solve optimization problem~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective functional in time.
The minimization problem is then reformulated as a (non)linear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, spectral  methods, and genetic algorithms.  
In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
following steps: (1) numerically discretize the FOM ODE (and hence the
\textit{integrand} of the objective function in problem~\eqref{eq:obj_gen_slab}) and 
(2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
To this end, we define time grids
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$ \EP{Shouldn't this be subset? We can't recover the continuous time interval},
$n=1,\ldots,\nslabs$ that
satisfy 
$\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
 = \timeEndArg{n}$ and, for notational purposes, we define $\timeWindowArg{n}{i} = \timeWindowArg{n-1}{\nstepsArg{n-1}+i}$ for $i < 0$ and $n=2,\ldots,\nslabs$.
%\KTC{use cdots, not
%hdots or ldots. Don't use $\defeq$ for the first and last times; these are
%just equalities, i.e., you happen to set them equal to those values, but they
%aren't identically the same:}
Figure~\ref{fig:slab_fig2} depicts such a discretization.
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-integration methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 4.5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 2$ for all $n$.} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.

@@

Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ (with $\lmsWidthArg{1}{1} = 1$) denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window. 

@@

&\residLMSWArg{n}{i} (\stateFOMDiscreteLMSArg{n}{i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
&\stateFOMDiscreteLMSArg{n}{i} = 
\begin{cases}
\stateFOMDiscreteLMSArg{n-1}{\nstepsArg{n-1}+i} & n = 2,\ldots,\nslabs, \; i \le 0 \\
\stateFOMIC & n=1, \; i=0,\end{cases}

@@

where $\stateFOMDiscreteLMSArg{n}{i} (\approx

@@

\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindowArg{n}{i-j}) \\

@@

\methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
$\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
$n=1,\ldots,\nslabs$ that satisfy
%	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
%	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
%	this should be the discrete solution, not the exact solution evaluated at
%	the final time.} \EP{You are correct, I think there is an error now in the minimize term though. It was $\in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}$, it should be $\trialspace\otimes \RR{\nstepsArg{n}+1}$}
\begin{align*}
	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
\objectiveArgLMS{n} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}). 
%%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
%%\begin{cases} \basisspaceArg{n} \spatialICDiscrete + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
%\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} 
\end{align*}
%Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
%\EP{Had to change the definition of the objective function to reflect stepping back into previous window}
In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
\begin{equation*}%\label{eq:obj_lms}
\begin{split} 
\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,1},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i}}), \\
&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
\RR{}_+, 
\end{split}
\end{equation*}
where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights, $\stateyDiscreteArgnt{n,i}$ for $i>0$ are the arguments, and for $i \le 0$
\EP{I think this is a little unclear, could include ICs in the definition of the objective function and split the arguments with a ;  Thoughts?} 
\begin{equation}\label{eq:wls_spatial_direct_args}
\stateyDiscreteArgnt{n,i}  \equiv 
\begin{cases}
\projectorSpatialArg{n}(\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i})  & n = 2,\ldots,\nslabs,  \\
\projectorSpatialArg{1}(\stateFOMIC)  & n = 1  \\
%\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
%\basisspaceArg{n} \basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 1.
\end{cases}
\end{equation}
with 
$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
Note that the boundary conditions are automatically satisfied by Eq.~\eqref{eq:wls_spatial_direct_args}.

%where 
%\begin{align*}
%\residArg{n,i} &: (\stateyArgnt{n,i},\ldots,\stateyArgnt{n,i-k(i)}) \mapsto  \sum_{j=0}^k \alpha_j \stateyArgnt{n,i-j} = \Delta t^{n,i} \sum_{j=0}^k \beta_j \velocity(\stateyArgnt{n,i-j}),
%\\
%               &: \RR{\fomdim} \otimes \RR{k(i)} \rightarrow \RR{\fomdim}, 
%\end{align*} 
%is the discrete residual resulting from the linear multistep scheme and 
%\methodAcronym\ solves the minimization problem over each time window, 
%\begin{align*}
%&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
%& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
%\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
%\approxstateIC & n=1. \end{cases} \end{align*}
	Equivalently, \methodAcronym\ sequentially computes the generalized
	coordinates
	$\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}$,
	$n=1,\ldots,\nslabs$ 
with $\genstateDiscreteArg{n,i}(\approx
	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdim}$
	that satisfy
\begin{equation}\label{eq:obj_gen_lms_final}
\begin{split}
	&
	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
%& \text{subject to} \hspace{0.2 in}
%\genstateyDiscreteArg{0} =
%\begin{cases} \genspatialICDiscrete  & n = 2,\ldots,\nslabs \\
%\genstateICOne& n=1. \end{cases} 
\end{split}
\end{equation}
%\begin{equation}\label{eq:obj_gen_lms_final}
%\begin{split}
%& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }
%\objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
%& \text{subject to} \hspace{0.2 in}
%\genstateDiscreteArg{n,0} =
%\begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
%\basisspace^T(\stateFOMIC - \stateIntercept)& n=1. \end{cases} 
%\end{split}
%\end{equation}
The optimization problem takes the form of a \textit{weighted least-squares
	problem}. 
We emphasize that optimization problem~\eqref{eq:obj_gen_lms_final} associates
	with an \spatialAcronym\ trial subspace characterized by a reduction in
	spatial complexity, but no reduction in temporal complexity.

@@

\begin{remark}
For the limiting case where $\nstepsArg{n} = 1$, $n=1,\ldots,\nslabs$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform 
quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
$\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
$\approxstateDiscreteArg{n,1}  = \approxstateLSPG^n$, $n=1,\ldots,\nslabs$ and  
\methodAcronym\ with \spatialAcronym\ trial
	subspaces solved via the direct approach recovers the LSPG approach. \KTC{Not exactly. To
	make this LSPG as presented earlier, also require the time grid here to
	align with the time grid used for time integration earlier. Also, this
	formulation requires allowing the stencil to go back to time steps at
	previous windows.}  
\end{remark} 
\subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
	\KTC{If the dynamics are linear, doesn't this just yield a linear
	least-squares problem that can be solved directly? We need to present this
	more clearly.}
Discretization through linear multistep methods (as well as other techniques) 
results in a least-squares problem. In particular, for the case where the full-order model~\eqref{eq:FOM} is nonlinear, the direct approach yields a nonlinear least-squares problem. 
A variety of algorithms exist for solving nonlinear least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method. The numerical experiments presented in this work comprise nonlinear systems and are solved via the Gauss--Newton method, and as such we outline this approach here.

To describe the Gauss--Newton method, we first define a ``vectorization" function 
\begin{align*}
 \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm}.
\end{align*}
The vectorized generalized coordinates over the $n$th time window are then
defined as \KTC{notationally, why are we doing a double bar instead of a bar?}
\EP{The notation for space--time in 2.4 uses bars. I don't want to confuse these generalized coordinates with those, they conceptually two different things} 
\begin{equation*}
\genstatecollocMatSlabArg{n} \defeq 
\unroll (\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
\end{equation*}
We now define the weighted space--time residual over the entire window as
\KTC{To make this valid for seeping back into the previous time window, the
second argument could be omitted here (okay because there is already a
superscript on the residual) and arbitrary previous solutions could be
included in the first residual here.}
\begin{equation*}
\residLMSSlabArg{n} : \genstatecollocMatySlabArg{} \mapsto \begin{bmatrix}
 \sqrt{\frac{\quadWeightsLMSScalarArg{n}{1}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n}; 
%\residLMSArg{n,0}( \basisspaceArg{n} \genstateyDiscreteArgnt{0} + \stateInterceptArg{n}),
\projectorSpatialArg{n}( \approxstateDiscreteArg{n,0}),
\ldots , \projectorSpatialArg{n}( \approxstateDiscreteArg{n,1 - \lmsWidthArg{n}{1} } )  ) \\
\vdots \\
 \sqrt{\frac{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}}}{2}} \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}; \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}-1} + \stateInterceptArg{n} ,  \ldots,\projectorSpatialArg{n}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - \lmsWidthArg{n}{\nstepsArg{n}} } + \stateInterceptArg{n} )) \\
\end{bmatrix},
\end{equation*}
where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
It is worth noting that, by design, 
\begin{equation*}
\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
=
\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n}) \bigg].
\end{equation*} 
Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method. The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and form the normal equations, and (3) solve the normal equations and update the state.

>>>>

Direct approaches solve optimization problem~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective functional in time.
The minimization problem is then reformulated as a (non)linear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, spectral  methods, and genetic algorithms.  
In the context of \spatialAcronym\ trial subspaces, the most straightforward direct solution approach consists of the 
following steps: (1) numerically discretize the FOM ODE (and hence the
\textit{integrand} of the objective function in problem~\eqref{eq:obj_gen_slab}) and 
(2) select a numerical quadrature rule to evaluate the \textit{integral} in~\eqref{eq:obj_gen_slab}.
To this end, we define time grids
$\{\timeWindowArg{n}{i}\}_{i=0}^{\nstepsArg{n}}\subseteq[\timeStartArg{n},\timeEndArg{n}]$ \EP{Shouldn't this be subset? We can't recover the continuous time interval},
$n=1,\ldots,\nslabs$ that
satisfy 
$\timeStartArg{n}=\timeWindowArg{n}{0} \le \cdots \le \timeWindowArg{n}{\nstepsArg{n}} 
 = \timeEndArg{n}$. %\KTC{use cdots, not
%hdots or ldots. Don't use $\defeq$ for the first and last times; these are
%just equalities, i.e., you happen to set them equal to those values, but they
%aren't identically the same:}
Figure~\ref{fig:slab_fig2} depicts such a discretization.
For the purposes of indexing between different windows, we additionally define a function:
$$\indexMapper: (n,i) \mapsto 
\begin{cases}
1,0 & n=1, \; i = 0,\\
n,i & n \ge 1, \; i > 0, \\
\indexMapper(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
\end{cases}$$
We now outline the direct solution approach for linear-multistep schemes; we note that the formulation for
other time-integration methods (e.g., Runge--Kutta) follows closely. 
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 4.5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid_timesteps.pdf} 
	\caption{Depiction of the $\nstepsArg{n}+1$ time instances over each window. In the figure, $\nstepsArg{n} = 2$ for all $n$.} 
\label{fig:slab_fig2} 
\end{centering} 
\end{figure}
%We now outline several commonly employed temporal discretization 
%schemes, specifically linear multistep and collocation (Runge--Kutta) methods, that leverage this time-grid to discretize the FOM ODE and 
%objective functional.

@@

Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $\lmsWidthArg{n}{i}$ time instances, where $\lmsWidthArg{n}{i}$ denotes the number of time steps employed by the scheme at the $i$th time instance over the $n$th window. 

@@

&\residLMSWArg{n}{i} (\stateFOMDiscreteArg{n,i};\stateFOMDiscreteLMSArg{n}{i-1},\ldots,\stateFOMDiscreteLMSArg{n}{i-\lmsWidthArg{n}{i}}) = \bz, \qquad i=1,\ldots,\nstepsArg{n}, \\
&\stateFOMDiscreteArg{1,0} = \stateFOMIC. 
%\begin{cases}
%\stateFOMDiscreteLMSArg{n-1}{0} & n = 2,\ldots,\nslabs \\

@@

where $\stateFOMDiscreteArg{n,i} (\approx

@@

\residLMSWArg{n}{i} &: (\stateyDiscreteArgnt{i};\stateyDiscreteArgnt{i-1},\ldots,\stateyDiscreteArgnt{i-\lmsWidthArg{n}{i}}) \mapsto  \frac{1}{\Delta t^{n,i}} \sum_{j=0}^{\lmsWidthArg{n}{i}} \alpha^{n,i}_j \stateyDiscreteArgnt{i-j} -  \sum_{j=0}^{\lmsWidthArg{n}{i}} \beta^{n,i}_j \velocity(\stateyDiscreteArgnt{i-j},\timeWindow^{\indexMapper(n,i-j)}) \\

@@

\methodAcronym\ with the direct approach and a linear multistep method sequentially computes the solutions
$\approxstateDiscreteArg{n,1},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$,
$n=1,\ldots,\nslabs$, that satisfy
%	\KTC{What is $\genstate^{n-1}(\timeEndArg{n-1} ) $ doing here? Two problems
%	(1) there is no notion of a generalized coordinate $\genstate$ yet, and (2)
%	this should be the discrete solution, not the exact solution evaluated at
%	the final time.} \EP{You are correct, I think there is an error now in the minimize term though. It was $\in \trialspace\otimes \RR{\lmsWidthArg{n}{i}+1}$, it should be $\trialspace\otimes \RR{\nstepsArg{n}+1}$}
\begin{align*}
	&\underset{(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \in \trialspace\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
\objectiveArgLMS{n} (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}). 
%%& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{0} =
%%\begin{cases} \basisspaceArg{n} \spatialICDiscrete + \stateInterceptArg{n}& n = 2,\ldots,\nslabs,\\
%\basisspaceArg{1} \genstateICOne  + \stateInterceptArg{1}& n=1. \end{cases} 
\end{align*}
%Employing a linear multistep method allows the objective \textit{functional}~\eqref{eq:obj} to be replaced with the objective \textit{function}
%\EP{Had to change the definition of the objective function to reflect stepping back into previous window}
In the above, $\objectiveArgLMS{n}$ is the discrete objective function at the $n$th time window and is defined as
\begin{equation*}%\label{eq:obj_lms}
\begin{split} 
\objectiveArgLMS{n} &\vcentcolon (\stateyDiscreteArgnt{n,1},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \mapsto
\frac{1}{2} \sum_{i=1}^{\nstepsArg{n}} \quadWeightsLMSScalarArg{n}{i} [\residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i})})]^T  \stweightingMatArgt{n}{\timeWindowArg{n}{i}} \residLMSWArg{n}{i}(\stateyDiscreteArgnt{n,i};\stateyDiscreteArgnt{n,i-1},\ldots, \stateyDiscreteArgnt{n,i-\lmsWidthArg{n}{i}}), \\
&\vcentcolon \RR{\fomdim} \otimes \RR{\nstepsArg{n}} \rightarrow
\RR{}_+, 
\end{split}
\end{equation*}
where $\quadWeightsLMSScalarArg{n}{i} \in \RRplus$ are quadrature weights, $\stateyDiscreteArgnt{n,i}$ for $i>0$ are the arguments, and for $i \le 0$
\EP{I think this is a little unclear, could include ICs in the definition of the objective function and split the arguments with a ;  Thoughts?} 
\begin{equation}\label{eq:wls_spatial_direct_args}
\stateyDiscreteArgnt{n,i}  \equiv 
\begin{cases}
\projectorSpatialArg{n}(\approxstateDiscreteArg{\indexMapper(n,i)})  & n = 2,\ldots,\nslabs,  \\
\projectorSpatialArg{1}(\stateFOMIC)  & n = 1, \\
%\basisspaceArg{n} \basisspaceTArg{n} (\approxstateDiscreteArg{n-1,\nstepsArg{n-1}+i} - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 2,\ldots,\nslabs,  \\
%\basisspaceArg{n} \basisspaceTArg{n} (\stateFOMIC - \stateInterceptArg{n}) + \stateInterceptArg{n} & n = 1.
\end{cases}
\end{equation}
where $\projectorSpatialArg{n} : \RR{\fomdim} \rightarrow \trialspaceArg{n}$ is the $\elltwo$-orthogonal projector and is defined as 
$$\projectorSpatialArg{n} : \stateyDiscreteArgnt{} \mapsto  \basisspaceArg{n}\basisspaceTArg{n} (\stateyDiscreteArgnt{} - \stateInterceptArg{n}) + \stateInterceptArg{n} .$$
Note that the boundary conditions are automatically satisfied by Eq.~\eqref{eq:wls_spatial_direct_args}.

%where 
%\begin{align*}
%\residArg{n,i} &: (\stateyArgnt{n,i},\ldots,\stateyArgnt{n,i-k(i)}) \mapsto  \sum_{j=0}^k \alpha_j \stateyArgnt{n,i-j} = \Delta t^{n,i} \sum_{j=0}^k \beta_j \velocity(\stateyArgnt{n,i-j}),
%\\
%               &: \RR{\fomdim} \otimes \RR{k(i)} \rightarrow \RR{\fomdim}, 
%\end{align*} 
%is the discrete residual resulting from the linear multistep scheme and 
%\methodAcronym\ solves the minimization problem over each time window, 
%\begin{align*}
%&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
%& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
%\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
%\approxstateIC & n=1. \end{cases} \end{align*}
	Equivalently, \methodAcronym\ sequentially computes the generalized
	coordinates
	$\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}$,
	$n=1,\ldots,\nslabs$ 
with $\genstateDiscreteArg{n,i}(\approx
	\genstateArgnt{n}(\timeWindowArg{n}{i}))\in\RR{\romdimArg{n}}$
	that satisfy
\begin{equation}\label{eq:obj_gen_lms_final}
\begin{split}
	&
	\underset{(\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}})\in\RR{\romdimArg{n}}\otimes \RR{\nstepsArg{n}}}{\text{minimize } }
\objectiveArgLMS{n} (\basisspaceArg{n} \genstateyDiscreteArg{1} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateyDiscreteArg{\nstepsArg{n}} + \stateInterceptArg{n}). 
%& \text{subject to} \hspace{0.2 in}
%\genstateyDiscreteArg{0} =
%\begin{cases} \genspatialICDiscrete  & n = 2,\ldots,\nslabs \\
%\genstateICOne& n=1. \end{cases} 
\end{split}
\end{equation}
%\begin{equation}\label{eq:obj_gen_lms_final}
%\begin{split}
%& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }
%\objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
%& \text{subject to} \hspace{0.2 in}
%\genstateDiscreteArg{n,0} =
%\begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
%\basisspace^T(\stateFOMIC - \stateIntercept)& n=1. \end{cases} 
%\end{split}
%\end{equation}
The optimization problem takes the form of a \textit{weighted least-squares
	problem}. 
We emphasize that optimization problem~\eqref{eq:obj_gen_lms_final} associates
	with an \spatialAcronym\ trial subspace characterized by a reduction in
	spatial complexity, but no reduction in temporal complexity.

@@

\begin{remark}
For the limiting case where $\nstepsArg{n} = 1$, $n=1,\ldots,\nslabs$ such that the window size is equivalent to the time step (i.e., $\DeltaSlabArg{n} = \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$), uniform 
quadrature weights are used, a uniform trial space is employed (i.e., $\trialspaceArg{n} = \trialspace, n=1,\ldots,\nslabs$), the weighting matrices are taken to be
$\stweightingMatOneArg{n} = \lspgWeightingArg{}, n=1,\ldots,\nslabs$, and the time instances $\timeWindowArg{n}{1} = t^n$, $n=1,\ldots,\nslabs$, then 
$\approxstateDiscreteArg{n,1}  = \approxstateLSPG^n$, $n=1,\ldots,\nslabs$ and  
\methodAcronym\ with \spatialAcronym\ trial
	subspaces solved via the direct approach recovers the LSPG approach. 
\end{remark} 
\subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
	\KTC{If the dynamics are linear, doesn't this just yield a linear
	least-squares problem that can be solved directly? We need to present this
	more clearly.}
Discretization through linear multistep methods (as well as other techniques) 
results in a least-squares problem. In particular, for the case where the full-order model~\eqref{eq:FOM} is nonlinear, the direct approach yields a nonlinear least-squares problem. 
A variety of algorithms exist for solving nonlinear least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method. The numerical experiments presented in this work comprise nonlinear systems and are solved via the Gauss--Newton method, and as such we outline this approach here.
